---
title: "Clean Data"
author: "Scott McKean"
output:
  pdf_document: default
  html_document: default
---

# Geoscience BC Study Code - Machine Learning Wrapper

```{r setup, include=FALSE}
library(geosciencebc2019008)
run_date = Sys.Date() %>% str_replace_all(., "-", "_")
knitr::opts_chunk$set(eval = FALSE)
```

# Load prepared data

This chunk loads our prepared data (see `prepare_data.Rmd`)

```{R}
saf_wells_sf <- read_rds('../wcfd_data/wcfd_saf_well_data.rds')
wcfd_data = TRUE
```

## ML Dataframe Prep

This chunk prepares our dataframe for the data science workflows (exploratory
data analysis, feature selection, model generation, and model interpretation). 
It specifies auxillary columns (those useful for information), predictor columns, 
and our target column(s), which for classificatio is the seimogenic (T/F)
association. Some of the columns, filtering, and imputation are specific for the 
classification problem, although the bits of code are very similar.

```{r}
aux_cols <- c(
  "unique_surv_id", "wa_num", "ground_elevtn",
  "mean_easting", "mean_northing", "on_prod_date"
  )

class_comp_feats = c(
  "calc_total_fluid_m3", "mean_rate_m3_min", "mean_proppant_per_stage_t",
  "proppant_cat", "horiz_wells_in_5km", "min_midpoint_dist",
  "calc_completed_length_m"
  )

class_geo_feats = c(
  "paleozoic_structure_mss", "geothermal_gradient_degc_km", "shmin_grasby",
  "distance_listric_faults_berger_m", "distance_normal_faults_berger_m"
  )

targets <- c("seismogenic", "max_mag")

ml_df <- saf_wells_sf %>%
  st_drop_geometry() %>% 
  dplyr::select(
    all_of(aux_cols), all_of(class_comp_feats), all_of(class_geo_feats), all_of(targets)) %>%
  dplyr::mutate_if(is.logical, as.numeric) %>%
  dplyr::filter(
    shmin_grasby > 0,
    !is.na(mean_proppant_per_stage_t),
    is.finite(calc_completed_length_m),
    min_midpoint_dist > 0) %>%
  dplyr::mutate(seismogenic = as.factor(seismogenic))
  
# impute non-causal features
ml_df = impute(ml_df, cols = list(
  mean_rate_m3_min = imputeMean()
  ))$data

# one hot encode categorical variables
ml_df <- createDummyFeatures(
  ml_df, target = "seismogenic", cols = c("proppant_cat")) %>% 
  removeConstantFeatures(.) %>%
  dplyr::select(
    -proppant_cat.ceramic, -proppant_cat.resin.coated, 
    -proppant_cat.sand, -proppant_cat.unknown) %>%
  tidyr::drop_na()

write_csv(ml_df, paste0("../wcfd_data/","final_wcfd_mldf_class.csv"))
write_rds(ml_df, paste0("../wcfd_data/","final_wcfd_mldf_class.rds"))
```

# Write csv and rds for future use

```{r}
regr_comp_feats  = c(
  "calc_total_fluid_m3", "calc_total_proppant_t", "calc_completed_length_m",
  "n_stages", "min_midpoint_dist", "mean_rate_m3_min"
  )

regr_geo_feats = c(
  "pressure_depth_ratio_kpa_m", "top_montney_structure_mss", "third_order_residual_m",
  "geothermal_gradient_degc_km", "distance_listric_faults_berger_m",
  "distance_normal_faults_berger_m"
  )

targets <- c("seismogenic", "max_mag")

ml_df <- saf_wells_sf %>%
  st_drop_geometry() %>% 
  dplyr::select(
    all_of(aux_cols), all_of(regr_comp_feats), all_of(regr_geo_feats), all_of(targets)) %>%
  dplyr::mutate_if(is.logical, as.numeric) %>%
  dplyr::filter(
    min_midpoint_dist > 0,
    is.finite(calc_completed_length_m)) %>%
  dplyr::mutate(seismogenic = as.factor(seismogenic))
  
# impute non-causal features
ml_df = impute(ml_df, cols = list(
  mean_rate_m3_min = imputeMean()
  ))$data %>%
  tidyr::drop_na()

write_csv(ml_df, paste0("../wcfd_data/","final_wcfd_mldf_regr.csv"))
write_rds(ml_df, paste0("../wcfd_data/","final_wcfd_mldf_regr.rds"))
```