---
title: "Process Seismic Catalogs"
author: "Scott McKean"
output: html_document
---

```{r setup, include=FALSE}
library(geosciencebc2019008)
```

Load the montney extents to subset and filter the catalogue

```{R}
# Load montney extents
gbc_blanking_region <- 
  sf::st_read(dsn = '../input/shapefiles', layer = 'montney_extents') %>%
  st_transform(crs = 26910)
```

# Load Files

Within the Montney extents, we have 1,107 events in the inducedseismicity.ca catalogue,
3,149 events from the NRCAN OF8337 report, and 10,693 events from the unpublished NRCAN report.
This substantially increases our catlaogue breadth! 

```{r}
is_cat <- '../input/catalogs/AlbertaCompCat2019-05.txt' %>%
  read.csv(., sep = '', header = TRUE) %>%
  dplyr::mutate(date_time = ISOdatetime(yr, mo, dy, hh, mm, ss, tz = 'UTC')) %>%
  dplyr::filter(Mpf >= -5,eventType != 2) %>%
  dplyr::mutate(mw = Mpf) %>%
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326, remove = FALSE) %>%
  sf::st_transform(crs = 26910) %>%
  dplyr::mutate(depth = ifelse(Dep_err < 0, NA, Dep)) %>%
  dplyr::select(depth, mw, date_time) %>%
  cbind(., sf::st_coordinates(.)) %>%
  dplyr::filter(st_contains_properly(gbc_blanking_region, .,sparse = FALSE))

nrc2014_cat <- read_xlsx('../input/catalogs/nrcan_of8335_catalog.xlsx') %>%
  as.data.frame() %>%
  janitor::clean_names() %>%
  dplyr::mutate(date_time = lubridate::ymd_hms(
    str_c(str_remove(y_m_d, " UTC"), " ",
          str_remove(h_m_s, "1899-12-31 ")))) %>%
  sf::st_as_sf(coords = c('lon_e', 'lat_n'), crs = 4326, remove = FALSE) %>%
  sf::st_transform(crs = 26910) %>%
  dplyr::mutate(mw_cor = ifelse(is.na(mw), ml, mw)) %>%
  dplyr::mutate(mw_coder = ifelse(is.na(mw), 'ml', 'mw')) %>%
  dplyr::select(depth, mw = mw_cor, date_time) %>%
  cbind(., sf::st_coordinates(.)) %>%
  dplyr::filter(st_contains_properly(gbc_blanking_region, .,sparse = FALSE))

nrc2017_cat <- read_xlsx('../input/catalogs/nrcan_ofxxxx_catalog.xlsx') %>%
  as.data.frame() %>%
  janitor::clean_names() %>%
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326, remove = FALSE) %>%
  sf::st_transform(crs = 26910) %>%
  dplyr::select(depth, mw = mag, date_time = datetime) %>%
  cbind(., sf::st_coordinates(.)) %>%
  dplyr::filter(st_contains_properly(gbc_blanking_region, .,sparse = FALSE))

all_cats <- rbind(is_cat, nrc2014_cat, nrc2017_cat) %>%
  dplyr::arrange(., date_time) %>%
  dplyr::mutate(event_id = 1:nrow(.))
```

# Remove duplicate values and filter > magnitude of completeness

In order to remove duplicate events from the catalogue, a one way search is conducted on
a combined dataframe of all events, sorted by datetime. Any events occuring after the first
event, within 10 seconds and within 10 km of the original event are flagged as a duplicate.
A total of 2,298 unique events are flagged as duplicates and removed from the composite 
catalogue. 

We conduct a basic magnitude of completeness analysis using the maximum magnitude method
with a rice binning. This provides a maximum magnitude of completeness of 0.25 for the
catalogue, which might be aggressive but should provide the best results for our 
machine learning workflow. The 'Sturges' binning provided a MaxM of 1, and the 'Freedman'
method provides a max magnitude of 0.05. We consider the 'Rice' method to provide a reasonable
balance between the two. We remove events below the catalogue magnitude of completeness (0.25).

We also remove all events before the year 2000.

This leaves us with a catalogue of 9,892 events. 

```{r}
flag_is_cat_duplicates <- function(i, all_cats){

  filt_cat <- all_cats[seq(i,i+10),]
  dist_bool <- st_distance(filt_cat[1,], filt_cat, sparse = FALSE) < set_units(10000, 'm')
  
  x <- which(dist_bool &
          (filt_cat$date_time - filt_cat[1,]$date_time) > 0 &
          (filt_cat$date_time - filt_cat[1,]$date_time) < 60)
  
  ifelse(is_empty(x), NA, filt_cat[x,]$event_id)
}

is_cat_duplicates <- purrr::map(.x = 1:nrow(all_cats), 
    .f = flag_is_cat_duplicates,
    all_cats = all_cats)

# filter out duplicates (since we did a one way search)
dup_vec <- simplify(is_cat_duplicates[!is.na(is_cat_duplicates)]) %>% unique()

# final composite catalogue
final_catalogue <- all_cats %>%
  dplyr::filter(!event_id %in% dup_vec) %>%
  dplyr::filter(date_time > ymd('2000-01-01'))

# get catalogue ECDF
ecdf <- process_cat(final_catalogue$mw, method = 'rice')

final_catalogue <- final_catalogue %>%
  dplyr::filter(mw >= 0.25)

write.csv(final_catalogue %>% st_drop_geometry(), 
          '../input/catalogs/20190430_combined_catalogue.csv')
```

Test to make sure catalog looks correct 

```{r}
final_catalogue <- read_csv('../input/catalogs/20190430_combined_catalogue.csv')
filt_cat <- final_catalogue %>% dplyr::filter(event_id %in% final_catalogue$event_id)

p1 = ggplot(filt_cat) +
  geom_histogram(aes(x = mw), bins = 40) +
  labs(x = 'Magnitude', y = 'Count') +
  theme_minimal()

p2 = ggplot() +
  geom_sf(data = filt_cat) +
  geom_sf(data = gbc_blanking_region, fill = NA) +
  labs(x = 'Easting', y = 'Northing', title = 'Final Catalogue') +
  coord_sf() +
  theme_minimal()
```
